---
title: "Axelar EVM Growth"
author: "Charliemarketplace"
date: "`r Sys.Date()`"
output:
  html_document:
    css: "styles.css"
    includes:
      in_header: header.html
    code_folding: hide
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

```{r, warning=FALSE, message=FALSE}
library(shroomDK)
library(reactable)
library(plotly)
library(visNetwork)
library(dplyr)

source("0_data_formatting.R")

```

# Intro

Axelar is a cross-chain transfer and messaging protocol and a leader in EVM <-> Cosmos 
transfers of tokens like USDC. Here, we focus on transfers via General Message Passing (GMP), 
most specifically the *Squid* Router.

Transaction originating from Ethereum Virtual Machine (EVM) blockchains that use the chain's 
local Squid Router address `0xce16f69375520ab01377ce7b88f5ba8c48f8d666` on each of:
Arbitrum, Avalanche, Binance Smart Chain, Ethereum Mainnet, and Polygon to send messages (most often USDC or axlUSDC) to other chains including each other but also Osmosis, Fantom, Moonbeam, etc. 

# Data 

On each `source chain` of: Arbitrum, Avalanche, Binance Smart Chain, Ethereum Mainnet, and Polygon  
a transaction is originated by an Externally Owned Account (EOA), i.e., a person (or possibly a bot). 

This `Origin From Address` is considered the unique, cross-EVM person address. 

Transactions from these EOA to the Squid Router where the Squid Router then *burns* 
the tokens (address: 0x00000...0000) are considered relevant GMP squid transfers. 
This is because axltokens on non-ETH chains are backed by tokens on ETH mainnet's Axelar Gateway. 
So these tokens can be burned and minted across non-ETH chains.

For Ethereum mainnet, instead of the burn address, Squid -> Gateway (`0x4f4495243837681061c4743b74b3eedf548d56a5`) 
is used.

```
with squid_to_burn AS (
SELECT 
BLOCK_NUMBER, BLOCK_TIMESTAMP, TX_HASH,
ORIGIN_FROM_ADDRESS as EOA, 
CONTRACT_ADDRESS as token_address,
event_inputs:value as raw_amount
 FROM 
 -- swap out each chain's event logs 
bsc.core.fact_event_logs
WHERE 
EVENT_NAME = 'Transfer' 
-- Squid Router burns token
AND EVENT_INPUTS:from = '0xce16f69375520ab01377ce7b88f5ba8c48f8d666'
AND EVENT_INPUTS:to = '0x0000000000000000000000000000000000000000'
),
```

Combining relevant data for each chain results in the `evm` table here where
raw amount is not decimal adjusted (e.g., USDC has 6 decimals).

```{r, warning = FALSE, message = FALSE}
reactable(
  head(evm)
)
```

Note: All data cut-off as of March 1, 2023 for reproducibility.

# Unique Users by source chain 

Ethereum, with the highest transaction fees, has the largest avg # of USDC 
sent in its transactions and the most $ total with the 2nd fewest transactions.

```{r, warning = FALSE, message = FALSE}
reactable(
  evm %>% group_by(sourcechain) %>% summarise(
    '# Unique Users' = length(unique(eoa)),
    '# Squid Tx' = length(unique(tx_hash)),
    'Avg Tx / User' = round(length(unique(tx_hash))/length(unique(eoa)),2),
    'Total $USDC Sent' = format(round(sum(raw_amount/1e6),0),big.mark = ","),
    'Avg $ Sent' = round(sum(raw_amount/1e6)/length(unique(tx_hash)),2)
  )
)
```

# User Level Visuals 

```{r, warning = FALSE, message = FALSE}
eoatx <- eoatbl %>% group_by(address) %>% 
  summarize(ntx = sum(amount),
            nsrc = length(unique(sourcechain)),
            ndest = length(unique(destinationchain))
              )

cdf_tx <- ecdf(eoatx$ntx)(seq(from = 0, to = max(eoatx$ntx), by = 1))
plot_ly(data = data.frame(), y = ~cdf_tx, type = 'scatter', mode = 'lines+markers') %>% 
  layout(xaxis = list(title = "Number of Transactions by User"),
         yaxis = list(title = "Cumulative % of EOAs"),
         title = list(
           text = "~96% of EVM GMP Users have 5 or less transactions", 
           y = 0.95)
  )


cdf_dest <- ecdf(eoatx$ndest)(seq(from = 0, to = max(eoatx$ndest), by = 1))
plot_ly(data = data.frame(), y = ~cdf_dest, type = 'scatter', mode = 'lines+markers') %>% 
  layout(xaxis = list(title = "# Destination Chains by User"),
         yaxis = list(title = "Cumulative % of EOAs"),
         title = list(
           text = "~95% of EVM GMP Users bridge to 2 or fewer chains", 
           y = 0.95)
  )

```

## User Level Network

```{r, warning = FALSE, message = FALSE}
# alphabetical order
src_chains <- sort(unique(evm$sourcechain))
dest_chains <- sort(unique(evm$destinationchain))

nodes <- data.frame(
  label = c(src_chains, dest_chains)
)

nodes$id <- 1:nrow(nodes)
nodes$group <- c(rep("source", length(src_chains)), rep("dest", length(dest_chains)))
nodes$shape <- c(rep("square", length(src_chains)), rep("circle", length(dest_chains)))
nodes$color <- c(rep("#fcdd42", length(src_chains)), rep("#e9e9f5", length(dest_chains)))
nodes$title <- paste0("<b>", nodes$label, "</b>")
  
nodes$y <- c(rep(-200, length(src_chains)), rep(200, length(dest_chains)))
nodes$x <- c(100*(length(src_chains):1) + 800, 100*(1:length(dest_chains)))

edges <- eoatbl

edges <- merge(
  edges, nodes[nodes$group == "source", c("id","label")], 
  by.x = "sourcechain", by.y = "label", all.x = TRUE
)

edges$from <- edges$id 
edges$id <- NULL
edges <- merge(
  edges, nodes[nodes$group == "dest", c("id","label")], 
  by.x = "destinationchain", by.y = "label", all.x = TRUE 
)
edges$to <- edges$id
edges$id <- NULL

edges$width <- floor(edges$amount/10)
edges$title <- paste0("EOA: ", edges$address,"<br>#GMP: ", edges$amount)

visNetwork(nodes, edges, background = "#FFFFFF",
           main = "5 EVM Source to all available Destinations",
           submain = "1 Line = 1 User; Width ~ # GMP by User") %>% 
    visOptions(highlightNearest = TRUE, nodesIdSelection = TRUE) %>% 
  visNodes(fixed = TRUE)
  
```


## Repeat-User Level Network 

Excluding those who have only done 1 transfer *ever*.

```{r, warning = FALSE, message = FALSE}
visNetwork(nodes, (edges[edges$address %in% eoatx$address[eoatx$ntx > 1], ]),
           background = "#FFFFFF",
            main = "Transfers among repeat GMP Users",
           submain = "1 Line = 1 User. Excludes EOAs with only 1 transfer ever") %>% 
    visOptions(highlightNearest = TRUE, nodesIdSelection = TRUE) %>% 
  visNodes(fixed = TRUE)

```


# Source-Destination Network

Aggregating up to 1 line per network-pair.

```{r, warning = FALSE, message = FALSE}

edges <- srcdest

edges <- merge(
  edges, nodes[nodes$group == "source", c("id","label")], 
  by.x = "sourcechain", by.y = "label", all.x = TRUE
)

edges$from <- edges$id 
edges$id <- NULL
edges <- merge(
  edges, nodes[nodes$group == "dest", c("id","label")], 
  by.x = "destinationchain", by.y = "label", all.x = TRUE 
)
edges$to <- edges$id
edges$id <- NULL

edges$width <- floor(log(edges$amount))
edges$title <- paste0("SRC: ", edges$sourcechain, 
                      "<br>Dest: ", edges$destinationchain,
                      "<br># GMP: ", edges$amount)


visNetwork(nodes, edges, 
           background = "#FFFFFF",
           main = "5 EVM Source to all available Destinations",
           submain = "1 Line = 1 Pair; Width ~ Log(#Transfers)") %>% 
    visOptions(highlightNearest = TRUE, nodesIdSelection = TRUE) %>% 
  visNodes(fixed = TRUE)
```


```{r, warning = FALSE, message = FALSE}
visNetwork(nodes, edges, 
           background = "#FFFFFF",
           main = "5 EVM Source to all available Destinations",
           submain = "1 Line = 1 Pair; Width ~ Log(#Transfers)") %>% 
    visOptions(highlightNearest = TRUE, nodesIdSelection = TRUE) %>% 
  visNodes(fixed = TRUE)

```

# Power Users

Looking at the subset of users that have done at least 5 GMP transfers originating on 
any of the 5 EVM chains of study. 

```{r, warning = FALSE, message = FALSE}
power_users <- eoatx[eoatx$ntx >= 5, ]
pucdf_tx <- ecdf(power_users$ntx)(sort(unique(power_users$ntx)))

plot_ly(data = data.frame(), x = sort(unique(power_users$ntx)), 
        y = ~pucdf_tx, type = 'scatter', mode = 'lines+markers') %>% 
  layout(xaxis = list(title = "Number of Transactions by User"),
         yaxis = list(title = "Cumulative % of EOAs"),
         title = list(
           text = "5 Addresses have used GMP >100 times \n 49 have used 6-99 times", 
           y = 0.95)
  )

```

# Power User Churn

```{r, warning = FALSE, message = FALSE}
evm_power <- evm[evm$eoa %in% power_users$address, ]
```

These `r length(unique(power_users$address))` most active addresses have done `r nrow(evm_power)` 
transactions ( `r round(nrow(evm_power)/nrow(evm)*100, 1)`% of the GMP transactions originating 
on the 5 main chains).

```{r, warning = FALSE, message = FALSE}

pu_summary <- evm_power %>% group_by(eoa) %>% 
  summarise(
    first_tx = min(block_timestamp),
    last_tx = max(block_timestamp),
    days_active = as.numeric(ceiling(difftime(max(block_timestamp), min(block_timestamp), units = "days"))),
    days_since = as.numeric(ceiling(difftime(as.Date("2023-03-01"), max(block_timestamp), units = "days"))),
    ntx = length(unique(tx_hash))
  ) %>% 
  mutate(tx_per_activeday = as.numeric(ntx)/as.numeric(days_active))

```

The earliest gmp transaction among power users was `r min(pu_summary$first_tx)` while 
the most recent in the data cutoff is `r max(pu_summary$last_tx)`. 

Looking at each users first GMP tx (any source chain) and last GMP tx (any source chain) 
identifies that of the users that have done at least 5 GMP transactions, they were Axelar users
for `r median(pu_summary$days_active)` days on median, and typically last used it `r median(pu_summary$days_since)` days ago.

```{r, warning = FALSE, message = FALSE}
plot_ly(pu_summary, x = ~days_active, type = "histogram", 
             xbins = list(size = 1)) %>% 
  layout(
  xaxis = list(title = "Days Active", showticklabels = TRUE),
  yaxis = list(title = "# of EOAs", showticklabels = TRUE),
  bargap = 0.2,
  title = list(text = "Median Power User is active 9 Days", y = 0.975)
)

plot_ly(pu_summary, x = ~days_since, type = "histogram", 
             xbins = list(size = 1)) %>% 
  layout(
  xaxis = list(title = "Days Since Last Tx", showticklabels = TRUE),
  yaxis = list(title = "# of EOAs", showticklabels = TRUE),
  bargap = 0.2,
  title = list(text = "Median Power User last active 10 days ago\n (Ref Date: 2023-03-01)", y = 0.975)
)

plot_ly(pu_summary, x = ~tx_per_activeday, type = "histogram", 
             xbins = list(size = 0.1)) %>% 
  layout(
  xaxis = list(title = "TX Per Active Day", showticklabels = TRUE),
  yaxis = list(title = "# of EOAs", showticklabels = TRUE),
  bargap = 0.2,
  title = list(text = "Median Power User did 1 TX per Active Day", y = 0.975)
)

```

Two key groups become apparent among power GMP users: 

- Persistent Early users who have a large # of transactions and have done transactions both in 2022 
and in February 2023.

- New active users that have started in 2023 and were active in the Feb 20th 2023 - March 1, 2023 time period.

```{r, warning = FALSE, message = FALSE}

plot_ly() %>% 
  add_trace(data = pu_summary, 
        x = ~first_tx, 
        y = ~last_tx,
        size = ~ntx, 
        type = "scatter", mode = "markers",
        name = "Short-term users") %>%
  add_trace(data = pu_summary %>% filter(first_tx < '2023-01-01' & last_tx >= '2023-02-01'), 
        x = ~first_tx, 
        y = ~last_tx,
        size = ~ntx, 
        type = "scatter", mode = "markers",
        name = "Persistent Early Users") %>% 
  add_trace(data = pu_summary %>% 
              filter(first_tx >= '2023-01-01' & last_tx >= '2023-02-20'),
        x = ~first_tx, 
        y = ~last_tx,
        size = ~ntx, 
        type = "scatter", mode = "markers",
        name = "New Active Users") %>%
  add_trace(data = pu_summary, 
            x = ~first_tx, y = ~first_tx, type = "scatter", mode = "lines", 
            line = list(color = "black", dash = 'dash', width = 2), name = "Same-Day Line") %>%
   layout(
  xaxis = list(title = "First TX", showticklabels = TRUE),
  yaxis = list(title = "Latest TX", showticklabels = TRUE),
  title = list(text = "Two Key Groups among Axelar Users", y = 0.975)
)

```

# Power User on-chain histories

Looking at the two groups and checking for all their activity across the 5 chains using
the Flipside Crypto `crosschain.core.address_tags` data.

```{r, warning = FALSE, message = FALSE}
query <- {
  "
    SELECT * FROM crosschain.core.address_tags
WHERE ADDRESS IN ('ADDRESSLIST') 
AND TAG_CREATED_AT <= '2023-03-01'
  "
}

# paste together the unique addresses to work within a SQL call. 
alist <- paste0(tolower(unique(pu_summary$eoa)), collapse = "','")

# swap parameters
query <- gsub('ADDRESSLIST', replacement = alist, x = query)

pu_history <- auto_paginate_query(query, api_key = readLines("api_key.txt"))  


pu_summary <- pu_summary %>% 
 mutate(
   group = case_when(
  first_tx < '2023-01-01' & last_tx >= '2023-02-01' ~ "Persistent Early User",
  first_tx >= '2023-01-01' & last_tx >= '2023-02-20' ~ "New Active User", 
  TRUE ~ "Other Short-Term User"
)
)

# The # of tag_name is not really important, just whether it is >0 or = 0
pu_actions <- pu_history %>% group_by(ADDRESS) %>% 
  summarise(
    n_nft_platforms = sum(TAG_NAME %in% c("opensea user", "looksrare user","x2y2 user", "rarible user")),
    top_nft_trader = sum(TAG_NAME %in% c("nft transactor top 5%", "nft transactor top 10%")) > 0, 
    top_balance = sum(TAG_NAME %in% c("eth top 1%", "wallet top 1%")) > 0,
    n_cex = sum(TAG_TYPE == 'cex'),
    n_active_chains = length(unique(BLOCKCHAIN))
  )

pu_actions <- merge(pu_actions, pu_summary[, c("eoa", "group")],
                    by.x = "ADDRESS", by.y = "eoa", all.x = TRUE)

```

A key differentiator between early persistent users and new-active users is the increased prevelance 
of trading NFTs on 1+ NFT platform. 

```{r, warning = FALSE, message = FALSE}
plot_ly(pu_actions) %>% 
  add_trace(data = pu_actions,
            x = ~n_nft_platforms,
        color = ~group,
        type = "histogram") %>% 
  layout(
    xaxis = list(title = "# ETH Mainnet NFT Platforms Used Ever"),
    yaxis = list(title = "# of Unique GMP Users (any evm chain)"),
    title = list(text = "Use of 1+ NFT Platforms small indication of GMP fit", y = 0.975)
  )

```

A sizeable proportion of power users have tried out 4+ EVM chains, while a rare few seem 
to have only used GMP from 1 of the 5 chains analyzed (e.g., only done ETH to Moonbeam).

```{r, warning = FALSE, message = FALSE}
plot_ly(pu_actions) %>% 
  add_trace(data = pu_actions,
            x = ~n_active_chains,
        color = ~group,
        type = "histogram") %>% 
  layout(
    title = list(text = "Use of at least 3 chains more indicative of GMP use", y = 0.975),
    xaxis = list(title = "# EVM Chains Used Ever"),
    yaxis = list(title = "# of Unique GMP Users (any evm chain)")
  )
```

Experimentation with multiple ETH mainnet Central Exchange transfers may be a useful indicator of GMP 
fitness.

```{r, warning = FALSE, message = FALSE}
plot_ly(pu_actions) %>% 
  add_trace(data = pu_actions,
            x = ~n_cex,
        color = ~group,
        type = "histogram") %>% 
  layout(
    title = list(text = "Use of 2+ ETH Mainnet CEX may be indicator of GMP fit", y = 0.975),
    xaxis = list(title = "# ETH Mainnet Central Exchanges Used Ever"),
    yaxis = list(title = "# of Unique GMP Users (any evm chain)")
  )

```

# Conclusion: EVM Market Fit / Opportunities

While Axelar GMP for EVM <-> EVM bridging still in its early days, there is a noticeable 
difference between Persistent Early Users from 2022 and newer more active users. 

Those most "sticky" to GMP (power users of 5+ GMP transactions across any combination of the 5 EVM chains 
reviewed) seem to:

- Experiment with NFTs (only a single Persistent Early User had tried even 1 NFT Platform: OpenSea, LooksRare, Rarible, or X2Y2).

- Use explicitly 3-6 EVM chains (early persistent users are all 2,3, or 5 and may be disproportionately arbitrage bots). 

- Used 2+ central exchange via ETH Mainnet, this may indicate users of GMP use their "main" KYC address as opposed to addresses they spin up that never touch a central exchange.

Using this criteria, let's identify all EVM addresses that may be a target addressable market for Axelar GMP
and see what chains they most often use. From there, alongside the provided network visualizations a plan for focusing 
on particular EVM chains & users can be generated. 


```{r, eval = FALSE}

# This query is not evaluated for speed reasons but a download is available in repo.
tam_query <- {
  "
SELECT ADDRESS, 
    SUM(CASE WHEN TAG_NAME IN ('opensea user', 'looksrare user', 'x2y2 user', 'rarible user') THEN 1 ELSE 0 END) AS n_nft_platforms,
    SUM(CASE WHEN TAG_TYPE = 'cex' THEN 1 ELSE 0 END) AS n_cex,
    COUNT(DISTINCT BLOCKCHAIN) AS n_active_chains,
    LISTAGG(DISTINCT BLOCKCHAIN, ',') WITHIN GROUP (ORDER BY BLOCKCHAIN) AS active_chains,
MAX(
    CASE 
        WHEN tag_type = 'activity' 
        THEN COALESCE(end_date, '2023-03-02') - 7
        ELSE COALESCE(end_date, start_date)
    END
) AS latest_active_date
FROM crosschain.core.address_tags
WHERE TAG_CREATED_AT <= '2023-03-02' 
GROUP BY ADDRESS
HAVING n_active_chains >= 3 AND n_cex >= 1 AND n_nft_platforms >= 1;
  "
}

tam_eoas <- auto_paginate_query(tam_query, api_key = readLines("api_key.txt"))

```

```{r, warning = FALSE, message = FALSE}
# read result of query for speed purposes 
tam_eoas <- read.csv("tam_eoas.csv", row.names = NULL, colClasses = 'character')
tam_eoas$LATEST_ACTIVE_DATE <- as.Date(tam_eoas$LATEST_ACTIVE_DATE, tz = 'UTC')
```

Of the `r nrow(tam_eoas)` EVM Addresses that meet the following criteria:

- Active on 3+ EVM chains *ever*
- At least 1 transaction from or to a central exchange
- At least 1 transaction on an NFT Marketplace 

`r sum(tam_eoas$LATEST_ACTIVE_DATE >= as.Date("2023-01-01"))` were active in 2023.

Of those active in 2023, the chain popularity order is ethereum (required to meet criteria), 
polygon, bsc, optimism, arbitrum, avalanche.

```{r, warning = FALSE, message = FALSE}
active_eoas <- tam_eoas[tam_eoas$LATEST_ACTIVE_DATE >= as.Date("2023-01-01"), ]

histories <- as.data.frame(table(unlist(strsplit(active_eoas$ACTIVE_CHAINS, split = ","))))
histories <- histories[order(histories$Freq, decreasing = TRUE), ]

colnames(histories) <- c("chain", "# Active TAM Users Ever")
reactable(histories)

```

Looking at a correlation matrix of which chains co-occur with each other among target addresses:

- Optimism & Arbitrum have significant co-occurrence 
- Avalanche and Binance have an enclave separate from Polygon/Optimism.

Because Axelar GMP does not currently support Optimism, interesting pairs of chains to focus on include: 

- Avalanche <-> BSC 
- Adding Optimism as an option for GMP and focusing on Arbitrum <-> Optimism.
- Arbitrum <-> Avalanche <-> Polygon

```{r, warning = FALSE, message = FALSE}
for(i in histories$chain){
active_eoas[i] <- as.numeric(grepl(i, active_eoas$ACTIVE_CHAINS))
}

corr <- round(cor(active_eoas[ , c("arbitrum", "avalanche", "bsc", "polygon", "optimism")]), 2)

plot_ly(z = corr, type = "heatmap", x = colnames(corr), y = colnames(corr), 
        colors = c("#d7191c", "#f7f7f7", "#2c7bb6"),
        colorscale = list(c(0, "#d7191c"), c(0.12, "#f7f7f7"), c(1, "#2c7bb6"))) %>% 
 colorbar(title = "Correlation", titleside = "top", 
                        tickmode = "array", tickvals = c(-1, 0, 1), ticktext = c("-1", "0", "1"), 
                        len = 0.5, thickness = 20, tickfont = list(size = 12), 
                        xpad = 10, ypad = 10, ticks = "inside") %>% 
  add_annotations(text = corr, x = rep(c(0,1,2,3,4), 5),
                  y = c(rep(0,5), rep(1,5), rep(2,5), rep(3,5), rep(4,5)),
                  showarrow = FALSE, font = list(color = "black", size = 20))
  
```

