---
title: "AxelScore Distributions"
author: "Charliemarketplace"
date: "`r Sys.Date()`"
output:
  html_document:
    css: "styles.css"
    includes:
      in_header: header.html
    code_folding: hide
    toc: true
    toc_float: true
editor_options: 
  chunk_output_type: console
---

# AxelScore Distributions

Before implementing a scoring methodology for Axelscore, this piece looks at historical 
Satellite and Squid transfers at the sender-source-destination-method level to 
assess current distributions of key stats at the sender level.

Note: In reality different senders from different source chains can be the same person. For the formal Axelscore app the user will link cross-chain sender addresses to get credit for multiple sender addresses.

Final scoring will take into account how these distributions would be collapsed by attributing multiple
senders to the same individual.

```{r, warning=FALSE, message=FALSE}

# library(shroomDK)
library(reactable)
library(plotly)
library(dplyr)

# Analysis is timestamped to prior to 2023-03-24 UTC
# https://flipsidecrypto.xyz/edit/queries/5278823a-aa1e-4677-910d-51a49cb7cda0

# NOT RUN 
axelscore_query <- {
  "
  
-- SELECT TOKENS ONLY!
with satellite_token_address_coingecko_id AS (
  SELECT COLUMN1 as TOKEN_ADDRESS, COLUMN2 as ID FROM (
VALUES ('uatom','cosmos'),('avalanche-uusdc','usd-coin'),('wavax-wei','avalanche-2'),
('uaxl','axelar'),('0x6e4e624106cb12e168e6533f8ec7c82263358940','axelar'),
('0x467719ad09025fcc6cf6f8311755809d45a5e5f3','axelar'),('0x44c784266cf024a60e8acf2427b9857ace194c5d','axelar'),
('0x8b1f4432f943c465a973fedc6d7aa50fc96f1f65','axelar'),('0x1b7c03bc2c25b8b5989f4bc2872cf9342cec80ae','axelar'),
('0x23ee2343b892b1bb63503a4fabc840e0e2c6810f','axelar'),('0x80d18b1c9ab0c9b5d6a6d5173575417457d00a12','cosmos'),
('0x33f8a5029264bcfb66e39157af3fea3e2a8a5067','cosmos'),('0x27292cf0016e5df1d8b37306b2a98588acbd6fca','cosmos'),
('0xddc9e2891fa11a4cc5c223145e8d14b44f3077c9','dai'),('0xc5fa5669e326da8b2c35540257cd48811f40a36b','dai'),
('0x4914886dbb8aad7a7456d471eaab10b06d42348d','frax'),('0x53adc464b488be8c5d7269b9abbce8ba74195c3a','frax'),
('0x750e4c4984a9e0f12978ea6742bc1c5d248f40ed','axlusdc'),('0xfab550568c688d5d8a52c7d794cb93edc26ec0ec','axlusdc'),
('0x4268b8f0b87b6eae5d897996e6b845ddbd99adf3','usd-coin'),('0xeb466342c4d449bc9f53a865d5cb90586f405215','usd-coin'),
('0xceed2671d8634e3ee65000edbbee66139b132fbf','tether'),('0xf976ba91b6bb3468c91e4f02e68b37bc64a57e66','tether'),
('0x7f5373ae26c3e8ffc4c77b7255df7ec1a9af52a6','tether'),('wbnb-wei','binancecoin'),('0x4fabb145d64652a948d72533023f6e7a623c7c53','binance-usd'),
('busd-wei','binance-usd'),('dai-wei','dai'),('0x6b175474e89094c44da98b954eedeac495271d0f','dai'),('dot-planck','polkadot'),
('weth-wei','ethereum'),('frax-wei','frax'),('0x853d955acef822db058eb8505911ed77f175b99e','frax'),('wftm-wei','fantom'),
('link-wei','chainlink'),('0x514910771af9ca656af840dff83e8264ecf986ca','chainlink'),('wmatic-wei','wmatic'),('mkr-wei','maker'),
('polygon-uusdc','usd-coin'),('uusdc','usd-coin'),('0xa0b86991c6218b36c1d19d4a2e9eb0ce3606eb48','usd-coin'),
('0x2791bca1f2de4661ed88a30c99a7a9449aa84174','usd-coin'),('0xb97ef9ef8734c71904d8002f8b6bc66dd9c48a6e','usd-coin'),('uusdt','tether'),
('0xdac17f958d2ee523a2206206994597c13d831ec7','tether'),('0xb31f66aa3c1e785363f0875a1b74e27b85fd66c7','avalanche-2'),
('0xbb4cdb9cbd36b01bd1cbaebf2de08d9173bc095c','binancecoin'),('wbtc-satoshi','bitcoin'),('0x2260fac5e5542a773aa44fbcfedf7c193bc2c599','bitcoin'),
('0xc02aaa39b223fe8d0a0e5c4f27ead9083c756cc2','ethereum'),('0x0d500b1d8e8ef31e21c99d1db9a6444d3adf1270','wmatic'))
),  

squid_hr_sender_og_dest_amount AS (
SELECT 'squid' as method, DATE_TRUNC('hour', BLOCK_TIMESTAMP) as hr, tx_hash, sender, 
token_address, token_symbol, amount, source_chain, destination_chain
FROM axelar.core.ez_squid
WHERE BLOCK_TIMESTAMP < '2023-03-24'
), 

satellite_hr_sender_og_dest_amount AS (
SELECT 'satellite' as method, DATE_TRUNC('hour', BLOCK_TIMESTAMP) as hr, tx_hash, sender, 
token_address, token_symbol, amount, source_chain, destination_chain
FROM axelar.core.ez_satellite
WHERE BLOCK_TIMESTAMP < '2023-03-24'
), 

all_sends AS 
(
SELECT * FROM squid_hr_sender_og_dest_amount 
UNION (SELECT * FROM satellite_hr_sender_og_dest_amount)
),

all_sends_labeled_id AS (
SELECT * FROM all_sends 
  INNER JOIN satellite_token_address_coingecko_id USING (TOKEN_ADDRESS)
),

-- Infill missing hour prices with most recent non-missing hour price
all_sends_priced AS (
SELECT *, 
coalesce(price, lag(price) IGNORE NULLS over (partition by ID order by HR)) as imputed_price
FROM all_sends_labeled_id
 LEFT JOIN (
    SELECT ID, DATE_TRUNC('hour', RECORDED_HOUR) as hr, 
    close as price
      FROM crosschain.core.fact_hourly_prices
    WHERE provider = 'coingecko'
  ) 
USING(ID,hr)
),

-- For transactions BEFORE fact_hourly_token_price has a price
-- Use the FIRST price ever recorded in fact_hourly_token_price
-- Otherwise, use actual hourly OR imputed (most recent) hourly price
-- Close -> price -> imputed_price -> final_price
all_sends_priced_final AS (
SELECT *, 
coalesce(imputed_price,
    FIRST_VALUE(imputed_price IGNORE NULLS) OVER (PARTITION BY ID ORDER BY HR)) AS final_price
FROM all_sends_priced
)

-- gotta send at least $1 between chains
SELECT SENDER, source_chain, destination_chain, method,
 count(*) as n_transfers, 
 sum(amount*final_price) as total_usd
FROM all_sends_priced_final
GROUP BY SENDER, source_chain, destination_chain, method
HAVING total_usd >= 1
  "
}

# NOT RUN 
# axelscore <- shroomDK::auto_paginate_query(axelscore_query, api_key = readLines("api_key.txt"))

```

# Data Prep

Creating a sender level table: 

```{r, warning=FALSE, message=FALSE}

axelscore <- read.csv("axlscore-selecttoken-sender-method-amounts.csv", row.names = NULL)
# fix avalanch -> avalanche
axelscore$SOURCE_CHAIN <- gsub(pattern = "$avalanch", 
                              replacement = "avalanche", 
                              x = axelscore$SOURCE_CHAIN)
axelscore$DESTINATION_CHAIN <- gsub(pattern = "$avalanch",
                                   replacement = "avalanche", 
                                   x = axelscore$DESTINATION_CHAIN)

evm <- c("ethereum","avalanche","polygon","binance","arbitrum")

# give cross_vm credit for nonevm -> evm OR evm -> nonevm 
axelscore <- axelscore %>% 
  mutate(
    cross_vm = ifelse(
      (SOURCE_CHAIN %in% evm & !(DESTINATION_CHAIN %in% evm)) |
      (!(SOURCE_CHAIN %in% evm) & DESTINATION_CHAIN %in% evm),
      1, 0
      ) 
  )

# Sender Summary:
#' Satellite total $USD 
#' Satellite total # Transfers 
#' Squid total $USD
#' Squid total # Transfers
#' Has done both squid & satellite (binary)
#' # of cross_vm transfers
#' $ of cross_vm transfers


sender_stats <- axelscore %>% group_by(SENDER) %>% 
  summarise(
    sat_usd = sum(TOTAL_USD[METHOD == "satellite"]),
    sat_transfers = sum(N_TRANSFERS[METHOD == "satellite"]),
    squid_usd = sum(TOTAL_USD[METHOD == "squid"]),
    squid_transfers = sum(N_TRANSFERS[METHOD == "squid"]),
    unique_methods = length(unique(METHOD)),
    crossvm_transfers = sum(N_TRANSFERS * cross_vm),
    crossvm_usd = sum(TOTAL_USD * cross_vm)
  )

reactable(head(sender_stats), 
          columns = list(
    SENDER = colDef(width = 80)),
    resizable = TRUE)

```

# Distributions 

## Total Satellite USD Value Transferred 

There are `r format(sum(sender_stats$sat_usd > 0), big.mark=",")` sender addresses 
that have transferred at least $1 of value via Satellite.

Among those that *have* transferred at least $1 via Satellite: 

```{r}
getsum <- function(stat){
  x = summary(stat)
  
  reactable(
    data.frame(
      stat = names(x),
      val = format(round(as.numeric(x),2), big.mark = ",")
    ),
    defaultColDef = colDef(width = 150, align ="right")
  )
}

get_distr <- function(df, x, 
                      binwidth = 30, 
                      color = "steelblue", 
                      title = "", 
                      xlab = "", 
                      ylab = "",
                      min_value = 0,
                      max_value = Inf) {
  
  df <- df %>% filter(get(x) <= max_value & get(x) >= min_value)
  bw <- binwidth
  bw_min <- min(df[[x]])
  bw_max <- max(df[[x]])
  breaks <- seq(bw_min, bw_max, bw)
  
  plot_ly(df, x = ~get(x), 
          type = "histogram",
          autobinx = FALSE,
                xbins = list(start = bw_min, end = bw_max, size = bw, 
                            autobin = FALSE, breaks = breaks),
               marker = list(color = color)) %>%
    layout(title = list(text = title), 
           xaxis = list(title = xlab), 
           yaxis = list(title = ylab),
           margin = list(l = 50, r = 50, t = 50, b = 50),
           showlegend = FALSE)
  
}

getsum(sender_stats$sat_usd[sender_stats$sat_usd > 0])

get_distr(sender_stats, 
          "sat_usd",
          binwidth = 100,
          color = "#3b82f680", 
          title = "Median Total Satellite usage is $311\n (Zooming in to <$20k)",
          xlab = "$USD Bucket", 
          ylab = "# of Senders", 
          min_value = 1,
          max_value = 20000)

```

## Total Satellite # Transfers

There are `r format(sum(sender_stats$sat_transfers > 0), big.mark=",")` sender addresses 
that have transferred at least once via Satellite.

Among those that have transferred at least once via Satellite: 

```{r}

getsum(sender_stats$sat_transfers[sender_stats$sat_transfers > 0])

get_distr(sender_stats, 
          "sat_transfers",
          binwidth = 1,
          color = "#3b82f680", 
          title = "Median # Satellite Transfers is 1\n (Zooming in to <50)",
          xlab = "# Transfers", 
          ylab = "# of Senders", 
          min_value = 1,
          max_value = 50)
```

This may be worth further investigation as only `r format(sum(sender_stats$sat_transfers > 1), big.mark=",")` senders have done 2+ transfers, implying only a 
`r round(100*sum(sender_stats$sat_transfers > 1)/sum(sender_stats$sat_transfers > 0),2)`% return rate (at the address level, recall 1 individual may have multiple sender addresses).

## Total Squid USD Value Transferred 


There are `r format(sum(sender_stats$squid_transfers > 0), big.mark=",")` sender addresses 
that have transferred at least once via Squid.

Among those that have transferred at least $1 via Squid: 

```{r}
getsum(sender_stats$squid_usd[sender_stats$squid_usd > 0])

get_distr(sender_stats, 
          "squid_usd",
          binwidth = 100,
          color = "#3b82f680", 
          title = "Median Total Squid usage is $20\n (Zooming in to <$5K)",
          xlab = "$USD Bucket", 
          ylab = "# of Senders", 
          min_value = 1,
          max_value = 5000)

```


## Total Squid # Transfers

There are `r format(sum(sender_stats$squid_transfers > 0), big.mark=",")` sender addresses 
that have transferred at least once via Squid.

Among those that have transferred at least once via Squid: 

```{r}
getsum(sender_stats$squid_transfers[sender_stats$squid_transfers > 0])

get_distr(sender_stats, 
          "squid_transfers",
          binwidth = 1,
          color = "#3b82f680", 
          title = "Median # Squid Transfers is 1\n (Zooming in to <50)",
          xlab = "# Transfers", 
          ylab = "# of Senders", 
          min_value = 1,
          max_value = 50)
```

This aligns to the previous [Axelar EVM Study](https://science.flipsidecrypto.xyz/axelar_evm_tam/) on Squid transfers. As only
`r round(100*sum(sender_stats$squid_transfers > 1)/sum(sender_stats$squid_transfers > 0),2)`%
of addresses returned to do 2+ transfers.

## USD Value by usage of one vs both methods 

Of the `r format(length(unique(sender_stats$SENDER)), big.mark = ',')` unique senders
not all are compatible with using both Satellite and Squid (Squid is not on Osmosis for 
example). 

When cross-chain identification in the final Axelscore product is available (i.e., 
when users bring their list of cross-chain addresses) we can better understand 
how usage between the two coincide. 

For now, we'll look at the `r format(sum(sender_stats$squid_transfers > 0), big.mark = ',')` Squid users and how key stats compare breaking apart the `r sum(sender_stats$squid_transfers > 0 & sender_stats$sat_transfers > 0)` that have used *both*.

```{r}
squids <- sender_stats %>% filter(squid_transfers > 0)

summaries <- squids %>% group_by(unique_methods) %>% 
  summarize(
    num_senders = n(),
    total_usd = sum(sat_usd + squid_usd),
    avg_usd = round(mean(sat_usd + squid_usd),0),
    median_usd = round(median(sat_usd + squid_usd),0),
    total_transfers = sum(sat_transfers + squid_transfers),
    avg_transfers = mean(sat_transfers + squid_transfers),
    median_transfers = median(sat_transfers + squid_transfers)
  ) %>% round(., 2) 

summaries$total_usd <- format(summaries$total_usd, big.mark = ",")
summaries$num_senders <- format(summaries$num_senders, big.mark = ",")
summaries$avg_usd <- format(summaries$avg_usd, big.mark = ",")
summaries$total_transfers <- format(summaries$total_transfers, big.mark = ",")

summaries <- t(summaries)
colnames(summaries) <- c("Squid Only Users", "Squid+Satellite Users")

  reactable(
   summaries,
    defaultColDef = colDef(width = 200, align ="right")
  )


```

## Cross Virtual Machine USD Value Transferred

While Squid transfers are exclusively within the Ethereum VM. Satellite transfers 
can be any combination including unrelated to EVM.

There are `r format(sum(sender_stats$crossvm_transfers > 0), big.mark = ',')` senders 
who initiated at least 1 crossvm transfer.

Among those that have transferred at least $1 across virtual machines (i.e., 
into EVM from outside or from EVM to outside): 

```{r}

getsum(sender_stats$crossvm_usd[sender_stats$crossvm_usd > 0])

get_distr(sender_stats, 
          "crossvm_usd",
          binwidth = 100,
          color = "#3b82f680", 
          title = "Median Total CrossVM USD Transferred in $367\n (Zooming in to <5000)",
          xlab = "$USD Bucket", 
          ylab = "# of Senders", 
          min_value = 1,
          max_value = 5000)
```

## Cross Virtual Machine Transfers

Among those that have transferred at least once across virtual machines (i.e., 
into EVM from outside or from EVM to outside): 

```{r}
getsum(sender_stats$crossvm_transfers[sender_stats$crossvm_transfers > 0])

get_distr(sender_stats, 
          "crossvm_transfers",
          binwidth = 1,
          color = "#3b82f680", 
          title = "Median # of CrossVM Transfers 1\n (Zooming in to <50)",
          xlab = "# Transfers", 
          ylab = "# of Senders", 
          min_value = 1,
          max_value = 50)

```

This may be worth further investigation as only `r format(sum(sender_stats$crossvm_transfers > 1), big.mark=",")` senders have done 2+ transfers, implying only a 
`r round(100*sum(sender_stats$crossvm_transfers > 1)/sum(sender_stats$crossvm_transfers > 0),2)`% return rate (at the address level, recall 1 individual may have multiple sender addresses).
